---
title: "Cpel models"
author: "Seoyeon Park B00768397"
date: "10/28/2021"
output:
  word_document: default
  html_document: default
---

Coccolithus pelagicus

#Scientific name is only one factor level so gam returns an error - need to remove s(scientific_name, bs = "re")
`
```{r}
load("data_subsets.RData")
load("data_crossval.RData")
```

```{r}
library(mgcv)
library(glmnet)
library(stats)
library(caret)
set.seed(101)
```


#MODEL G 

##Cpel_modG

```{r}
Cpel_modG <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc","tp")) , data = data.Cpel)
summary(Cpel_modG)
```

###train_cpel_modG

```{r}
train_cpel_modG <-  gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc","tp")) , data = train.Cpel)
summary(train_cpel_modG)
```

```{r}
plot(train_cpel_modG)
```

###test_cpel_modG

```{r}
test_cpel_modG <- train_cpel_modG %>% predict(test.Cpel, type = "response")
summary(test_cpel_modG)
```

RMSE of test data: 
```{r}
error_cpel_modG <- test.Cpel$cells_per_litre - test_cpel_modG
RMSE_cpel_modG <- sqrt(mean(error_cpel_modG^2))
print(RMSE_cpel_modG)
```

prediction error rate:
```{r}
RMSE_cpel_modG/mean(test.Cpel$cells_per_litre)
```

Actual vs predicted: 
```{r}
ggplot(data = test.Cpel, mapping = aes(x = cells_per_litre, y = test_cpel_modG)) + geom_point() + geom_abline(intercept = 0, slope = 1, color = "red", size = 1)
```

#MODEL GS - environmental variables + group smoother scientific_name

##Cpel_modGS_p

```{r}
Cpel_modGS_p <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(p_an, scientific_name, bs = "fs"), data = data.Cpel, method = "REML")
summary(Cpel_modGS_p)
```

###train_cpel_modGS_p

```{r}
train_cpel_modGS_p <-  gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(p_an, scientific_name, bs = "fs"), data = train.Cpel, method = "REML")
summary(train_cpel_modGS_p)
```

```{r}
plot(train_cpel_modGS_p)
```

###test_cpel_modGS_p

```{r}
test_cpel_modGS_p <- train_cpel_modGS_p %>% predict(test.Cpel, type = "response")
summary(test_cpel_modGS_p)
```

RMSE of test data: 
```{r}
error_cpel_modGS_p <- test.Cpel$cells_per_litre - test_cpel_modGS_p
RMSE_cpel_modGS_P <- sqrt(mean(error_cpel_modGS_p^2))
print(RMSE_cpel_modGS_P)
```

prediction error rate:
```{r}
RMSE_cpel_modGS_P/mean(test.Cpel$cells_per_litre)
```

Actual vs predicted: 
```{r}
ggplot(data = test.Cpel, mapping = aes(x = cells_per_litre, y = test_cpel_modGS_p)) + geom_point() + geom_abline(intercept = 0, slope = 1, color = "red", size = 1)
```

##Cpel_modGS_sistar

```{r}
Cpel_modGS_sistar <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(si_star, scientific_name, bs = "fs"), data = data.Cpel, method = "REML")
summary(Cpel_modGS_sistar)
```

###train_cpel_modGS_sistar

```{r}
train_cpel_modGS_sistar <-  gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(si_star, scientific_name, bs = "fs"), data = train.Cpel, method = "REML")
summary(train_cpel_modGS_sistar)
```

```{r}
plot(train_cpel_modGS_sistar)
```

###test_cpel_modGS_sistar

```{r}
test_cpel_modGS_sistar <- train_cpel_modGS_sistar %>% predict(test.Cpel, type = "response")
summary(test_cpel_modGS_sistar)
```

RMSE of test data: 
```{r}
error_cpel_modGS_sistar <- test.Cpel$cells_per_litre - test_cpel_modGS_sistar
RMSE_cpel_modGS_sistar <- sqrt(mean(error_cpel_modGS_sistar^2))
print(RMSE_cpel_modGS_sistar)
```

prediction error rate:
```{r}
RMSE_cpel_modGS_sistar/mean(test.Cpel$cells_per_litre)
```

Actual vs predicted: 
```{r}
ggplot(data = test.Cpel, mapping = aes(x = cells_per_litre, y = test_cpel_modGS_sistar)) + geom_point() + geom_abline(intercept = 0, slope = 1, color = "red", size = 1)
```

##Cpel_modGS_nistar

```{r}
Cpel_modGS_nistar <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(ni_star, scientific_name, bs = "fs"), data = data.Cpel, method = "REML")
summary(Cpel_modGS_nistar)
```


###train_cpel_modGS_nistar

```{r}
train_cpel_modGS_nistar <-  gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(ni_star, scientific_name, bs = "fs"), data = train.Cpel, method = "REML")
summary(train_cpel_modGS_nistar)
```

```{r}
plot(train_cpel_modGS_nistar)
```

###test_cpel_modGS_nistar

```{r}
test_cpel_modGS_nistar <- train_cpel_modGS_nistar %>% predict(test.Cpel, type = "response")
summary(test_cpel_modGS_nistar)
```

RMSE of test data: 
```{r}
error_cpel_modGS_nistar <- test.Cpel$cells_per_litre - test_cpel_modGS_nistar
RMSE_cpel_modGS_nistar <- sqrt(mean(error_cpel_modGS_nistar^2))
print(RMSE_cpel_modGS_nistar)
```

prediction error rate:
```{r}
RMSE_cpel_modGS_nistar/mean(test.Cpel$cells_per_litre)
```

Actual vs predicted: 
```{r}
ggplot(data = test.Cpel, mapping = aes(x = cells_per_litre, y = test_cpel_modGS_nistar)) + geom_point() + geom_abline(intercept = 0, slope = 1, color = "red", size = 1)
```

##Cpel_modGS_s

```{r}
Cpel_modGS_s <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(s_an, scientific_name, bs = "fs"), data = data.Cpel, method = "REML")
summary(Cpel_modGS_s)
```

###train_cpel_modGS_s

```{r}
train_cpel_modGS_s <-  gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(s_an, scientific_name, bs = "fs"), data = train.Cpel, method = "REML")
summary(train_cpel_modGS_s)
```

```{r}
plot(train_cpel_modGS_s)
```

###test_cpel_modGS_s

```{r}
test_cpel_modGS_s <- train_cpel_modGS_s %>% predict(test.Cpel, type = "response")
summary(test_cpel_modGS_s)
```

RMSE of test data: 
```{r}
error_cpel_modGS_s <- test.Cpel$cells_per_litre - test_cpel_modGS_s
RMSE_cpel_modGS_s <- sqrt(mean(error_cpel_modGS_s^2))
print(RMSE_cpel_modGS_s)
```

prediction error rate:
```{r}
RMSE_cpel_modGS_s/mean(test.Cpel$cells_per_litre)
```

Actual vs predicted: 
```{r}
ggplot(data = test.Cpel, mapping = aes(x = cells_per_litre, y = test_cpel_modGS_s)) + geom_point() + geom_abline(intercept = 0, slope = 1, color = "red", size = 1)
```

##Cpel_modGS_t

```{r}
Cpel_modGS_t <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(t_an, scientific_name, bs = "fs"), data = data.Cpel, method = "REML")
summary(Cpel_modGS_t)
```
###train_cpel_modGS_t

```{r}
train_cpel_modGS_t <-  gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(t_an, scientific_name, bs = "fs"), data = train.Cpel, method = "REML")
summary(train_cpel_modGS_t)
```

```{r}
plot(train_cpel_modGS_t)
```

###test_cpel_modGS_t

```{r}
test_cpel_modGS_t <- train_cpel_modGS_t %>% predict(test.Cpel, type = "response")
summary(test_cpel_modGS_t)
```

RMSE of test data: 
```{r}
error_cpel_modGS_t <- test.Cpel$cells_per_litre - test_cpel_modGS_t
RMSE_cpel_modGS_t <- sqrt(mean(error_cpel_modGS_t^2))
print(RMSE_cpel_modGS_t)
```

prediction error rate:
```{r}
RMSE_cpel_modGS_t/mean(test.Cpel$cells_per_litre)
```

Actual vs predicted: 
```{r}
ggplot(data = test.Cpel, mapping = aes(x = cells_per_litre, y = test_cpel_modGS_t)) + geom_point() + geom_abline(intercept = 0, slope = 1, color = "red", size = 1)
```
#MODEL GI - environmental variables + individual smoother scientific_name

##Cpel_modGI_p

```{r}
Cpel_modGI_p <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(p_an, by = scientific_name, bs = "fs", m = 1), data = data.Cpel, method = "REML")
summary(Cpel_modGI_p)
```

###train_cpel_modGI_p
```{r}
train_cpel_modGI_p <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(p_an, by = scientific_name, bs = "fs", m = 1), data = train.Cpel, method = "REML")
summary(train_cpel_modGI_p)
```

###test_cpel_modGI_p

```{r}
test_cpel_modGI_p <- train_cpel_modGI_p %>% predict(test.Cpel, type = "response")
summary(test_cpel_modGI_p)
```

RMSE of test data: 
```{r}
error_cpel_modGI_p <- test.Cpel$cells_per_litre - test_cpel_modGI_p
RMSE_cpel_modGI_p <- sqrt(mean(error_cpel_modGI_p^2))
print(RMSE_cpel_modGI_p)
```

prediction error rate:
```{r}
RMSE_cpel_modGI_p/mean(test.Cpel$cells_per_litre)
```

Actual vs predicted: 
```{r}
ggplot(data = test.Cpel, mapping = aes(x = cells_per_litre, y = test_cpel_modGI_p)) + geom_point() + geom_abline(intercept = 0, slope = 1, color = "red", size = 1)
```

##Cpel_modGI_sistar

```{r}
Cpel_modGI_sistar <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(si_star, by = scientific_name, bs = "fs", m = 1), data = data.Cpel, method = "REML")
summary(Cpel_modGI_sistar)
```

###train_cpel_modGI_sistar
```{r}
train_cpel_modGI_sistar <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(si_star, by = scientific_name, bs = "fs", m = 1), data = train.Cpel, method = "REML")
summary(train_cpel_modGI_sistar)
```

###test_cpel_modGI_sistar

```{r}
test_cpel_modGI_sistar <- train_cpel_modGI_sistar %>% predict(test.Cpel, type = "response")
summary(test_cpel_modGI_sistar)
```

RMSE of test data: 
```{r}
error_cpel_modGI_sistar <- test.Cpel$cells_per_litre - test_cpel_modGI_sistar
RMSE_cpel_modGI_sistar <- sqrt(mean(error_cpel_modGI_sistar^2))
print(RMSE_cpel_modGI_sistar)
```

prediction error rate:
```{r}
RMSE_cpel_modGI_sistar/mean(test.Cpel$cells_per_litre)
```

Actual vs predicted: 
```{r}
ggplot(data = test.Cpel, mapping = aes(x = cells_per_litre, y = test_cpel_modGI_sistar)) + geom_point() + geom_abline(intercept = 0, slope = 1, color = "red", size = 1)
```

##Cpel_modGI_nistar

```{r}
Cpel_modGI_nistar <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(ni_star, by = scientific_name, bs = "fs", m = 1), data = data.Cpel, method = "REML")
summary(Cpel_modGI_nistar)
```

###train_cpel_modGI_nistar
```{r}
train_cpel_modGI_nistar <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(ni_star, by = scientific_name, bs = "fs", m = 1), data = train.Cpel, method = "REML")
summary(train_cpel_modGI_nistar)
```

###test_cpel_modGI_nistar

```{r}
test_cpel_modGI_nistar <- train_cpel_modGI_nistar %>% predict(test.Cpel, type = "response")
summary(test_cpel_modGI_nistar)
```

RMSE of test data: 
```{r}
error_cpel_modGI_nistar <- test.Cpel$cells_per_litre - test_cpel_modGI_nistar
RMSE_cpel_modGI_nistar <- sqrt(mean(error_cpel_modGI_nistar^2))
print(RMSE_cpel_modGI_nistar)
```

prediction error rate:
```{r}
RMSE_cpel_modGI_nistar/mean(test.Cpel$cells_per_litre)
```

Actual vs predicted: 
```{r}
ggplot(data = test.Cpel, mapping = aes(x = cells_per_litre, y = test_cpel_modGI_nistar)) + geom_point() + geom_abline(intercept = 0, slope = 1, color = "red", size = 1)
```

##Cpel_modGI_s

```{r}
Cpel_modGI_s <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(s_an, by =  scientific_name, bs = "fs", m = 1), data = data.Cpel, method = "REML")
summary(Cpel_modGI_s)
```

###train_cpel_modGI_s
```{r}
train_cpel_modGI_s <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(s_an, by = scientific_name, bs = "fs", m = 1), data = train.Cpel, method = "REML")
summary(train_cpel_modGI_s)
```

###test_cpel_modGI_s

```{r}
test_cpel_modGI_s <- train_cpel_modGI_s %>% predict(test.Cpel, type = "response")
summary(test_cpel_modGI_s)
```

RMSE of test data: 
```{r}
error_cpel_modGI_s <- test.Cpel$cells_per_litre - test_cpel_modGI_s
RMSE_cpel_modGI_s <- sqrt(mean(error_cpel_modGI_s^2))
print(RMSE_cpel_modGI_s)
```

prediction error rate:
```{r}
RMSE_cpel_modGI_s/mean(test.Cpel$cells_per_litre)
```

Actual vs predicted: 
```{r}
ggplot(data = test.Cpel, mapping = aes(x = cells_per_litre, y = test_cpel_modGI_s)) + geom_point() + geom_abline(intercept = 0, slope = 1, color = "red", size = 1)
```

##Cpel_modGI_t

```{r}
Cpel_modGI_t <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(t_an, by = scientific_name, bs = "fs", m = 1), data = data.Cpel, method = "REML")
summary(Cpel_modGI_t)
```

###train_cpel_modGI_t
```{r}
train_cpel_modGI_t <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(t_an, by = scientific_name, bs = "fs", m = 1), data = train.Cpel, method = "REML")
summary(train_cpel_modGI_t)
```

###test_cpel_modGI_t

```{r}
test_cpel_modGI_t <- train_cpel_modGI_t %>% predict(test.Cpel, type = "response")
summary(test_cpel_modGI_t)
```

RMSE of test data: 
```{r}
error_cpel_modGI_t <- test.Cpel$cells_per_litre - test_cpel_modGI_t
RMSE_cpel_modGI_t <- sqrt(mean(error_cpel_modGI_t^2))
print(RMSE_cpel_modGI_t)
```

prediction error rate:
```{r}
RMSE_cpel_modGI_t/mean(test.Cpel$cells_per_litre)
```

#AIC values
```{r}
AIC(train_cpel_modG, train_cpel_modGS_p, train_cpel_modGS_sistar, train_cpel_modGS_nistar, train_cpel_modGS_s, train_cpel_modGS_t, train_cpel_modGI_p, train_cpel_modGI_sistar, train_cpel_modGI_nistar, train_cpel_modGI_s, train_cpel_modGI_t)
```







