---
title: "Data - Honours"
author: "Seoyeon Park B00768397"
date: "9/7/2021"
output: html_document
---

```{r}
rm(list = ls())
```

# Setting up
Load the Coccobase library, mgcv, and other required packages:
```{r}
library(coccobase)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(maps)
library(viridis)
library(kableExtra)
library(stringr)
library(gratia)
library(mgcv)
library(Metrics)
library("stringr")
```


Needed data:
```{r}
head(environment_data)
head(cocco)
```

Join environment data and cocco together
lat/lon is the 1st/4th column in environment_data - switch names to "latitude" and "longitude" to use left_join()

```{r}
environment_data <- environment_data %>% dplyr::rename(latitude = lat, longitude = lon)
```


```{r}
cocco <- cocco %>% dplyr::select(month,longitude,latitude, cells_per_litre, scientific_name, depth) %>% mutate(longitude = round((longitude-0.5),digits = 0) + 0.5) %>% mutate(latitude = round((latitude-0.5), digits = 0) + 0.5) %>% mutate(cells_per_litre = log(cells_per_litre)) %>% filter(cells_per_litre > 0) 

data <- left_join(cocco,environment_data, by = c("longitude","latitude","month"))
data
```  

Refit the models with salinity >=30, depth <=50, replace silicate with silicate - nitrate (si_star), replace nitrate with nitrate - 16*phosphate (ni_star), Set longitude and latitude to North Atlantic (longitude from -50 to 0, latitude > 40)
```{r}
data <- data %>% filter(s_an >= 30) %>% filter(depth <=50) %>% mutate(ni_star = n_an - 16*p_an) %>% mutate(si_star = i_an - n_an) %>% subset(latitude >= 40 & longitude <= 0 & longitude >= -50) %>% mutate(cells_per_litre = log(cells_per_litre))
data
```

summarizing the number species filtered in NA

```{r}
data %>% count(scientific_name) %>% arrange(-n)
```

Final data should only include the following species:
E.hux (n = 1367)
Coccolithus pelagicus  (n = 572)
Syracosphaera pulchra (n = 240)
total coccolithophore species (n = 5823) - includes all Coccolithophore observations in NA
Coccolithus species (n = 803) - includes Coccolithus pelagicus and Coccolithus spp. 
Syracosphaera species (n = 790) - includes all species with family name Syracosphaera   

Summary of environmental variables

```{r}
data %>% select(c(p_an, ni_star, si_star, s_an, t_an)) %>% summary()
```

Scientific name must be unordered
```{r}
data <- transform(data, scientific_name= factor(scientific_name, ordered=FALSE)) 
```

```{r}
data.full <- data
data.Ehux <- data %>% filter(scientific_name  == "Emiliania huxleyi")
data.Coccolithus <- data[str_detect(data$scientific_name, "Coccolithus "), ]
data.Cpel <- data %>% filter(scientific_name == "Coccolithus pelagicus")
data.Syracosphaera <- data[str_detect(data$scientific_name, "Syracosphaera"), ]
data.Spul <- data %>% filter(scientific_name == c("Syracosphaera pulchra"))
```

```{r}
save(data.full, data.Coccolithus, data.Cpel, data.Ehux, data.Spul, data.Syracosphaera, file = "data_subsets.RData")
```

























