---
title: "cross validation data"
author: "Seoyeon Park B00768397"
date: "11/16/2021"
output: html_document
---

```{r}
rm(list = ls())
```

```{r}
load("data_subsets.RData")
```

```{r}
set.seed(101)
```


Training and Testing (60:40 split)

Coccolithophore:

```{r}
n = nrow(data.full)
split = sample(1:n, size = round(0.6*n), replace = F)
train.full = data.full[split,]
test.full = data.full[-split,]
```

Training data has 3493 observations
Testing data has 2329 observations 
Total of 5822 observations

Coccolithus:

```{r}
n = nrow(data.Coccolithus)
split = sample(1:n, size = round(0.6*n), replace = F)
train.Coccolithus = data.Coccolithus[split,]
test.Coccolithus = data.Coccolithus[-split,]
```

Training data has 482 observations
Testing data has 321 observations
Total of 803 observations 

Syracosphaera:

```{r}
n = nrow(data.Syracosphaera)
split = sample(1:n, size = round(0.6*n), replace = F)
train.Syracosphaera = data.Syracosphaera[split,]
test.Syracosphaera = data.Syracosphaera[-split,]
```

Training data has 474 observations
Testing data has 316 observations
Total of 790 observations

Emiliania huxleyi:

```{r}
n = nrow(data.Ehux)
split = sample(1:n, size = round(0.6*n), replace = F)
train.Ehux = data.Ehux[split,]
test.Ehux = data.Ehux[-split,]
```

Training data has 820 observations
Testing data has 547 observations
Total of 1367 observations

Coccolithus pelagicus:

```{r}
n = nrow(data.Cpel)
split = sample(1:n, size = round(0.6*n), replace = F)
train.Cpel = data.Cpel[split,]
test.Cpel = data.Cpel[-split,]
```

Training data has 343 observations
Testing data has 229 observations
Total of 572 observations

Syracosphaera Pulchra:

```{r}
n = nrow(data.Spul)
split = sample(1:n, size = round(0.6*n), replace = F)
train.Spul = data.Spul[split,]
test.Spul = data.Spul[-split,]
```

Training data has 144 observations
Testing data has 96 observations
Total of 240 observations


```{r}
save(train.full, test.full, train.Ehux, test.Ehux, train.Coccolithus, test.Coccolithus, train.Cpel, test.Cpel, train.Syracosphaera, test.Syracosphaera, train.Spul, test.Spul, file = "data_crossval.RData")
```







