---
title: "Ehux models"
author: "Seoyeon Park B00768397"
date: "10/28/2021"
output:
  word_document: default
  html_document: default
---

#Scientific name is only one factor level so gam returns an error - need to remove s(scientific_name, bs = "re")
`
```{r}
load("data_subsets.RData")
```

```{r}
library(mgcv)
library(glmnet)
library(stats)
library(caret)
set.seed(101)
```


##MODEL G

#Ehux_modG

```{r}
Ehux_modG <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc","tp")) , data = data.Ehux)
summary(Ehux_modG)
```

##train_ehux_modG

```{r}
train_ehux_modG <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc","tp")) , data = train.Ehux)
summary(train_ehux_modG)
```

```{r}
plot(train_ehux_modG)
```

##test_ehux_modG

```{r}
test_ehux_modG <- train_ehux_modG %>% predict(test.Ehux, type = "response")
summary(test_ehux_modG)
```

RMSE of test data: 
```{r}
error_ehux_modG <- test.Ehux$cells_per_litre - test_ehux_modG
RMSE_ehux_modG <- sqrt(mean(error_ehux_modG^2))
print(RMSE_ehux_modG)
```

prediction error rate:
```{r}
RMSE_ehux_modG/mean(test.Ehux$cells_per_litre)
```

##MODEL GS - environmental variables + group smoother scientific_name

#Ehux_modGS_p

```{r}
Ehux_modGS_p <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(p_an, scientific_name, bs = "fs"), data = data.Ehux, method = "REML")
summary(Ehux_modGS_p)
```

##train_ehux_modGS_p
```{r}
train_ehux_modGS_p <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(p_an, scientific_name, bs = "fs"), data = train.Ehux, method = "REML")
summary(train_ehux_modGS_p)
```

##test_ehux_modGS_p

```{r}
test_ehux_modGS_p <- train_ehux_modGS_p %>% predict(test.Ehux, type = "response")
summary(test_ehux_modGS_p)
```

RMSE of test data: 
```{r}
error_ehux_modGS_p <- test.Ehux$cells_per_litre - test_ehux_modGS_p
RMSE_ehux_modGS_p <- sqrt(mean(error_ehux_modGS_p^2))
print(RMSE_ehux_modGS_p)
```

prediction error rate:
```{r}
RMSE_ehux_modGS_p/mean(test.Ehux$cells_per_litre)
```

#Ehux_modGS_sistar

```{r}
Ehux_modGS_sistar <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(si_star, scientific_name, bs = "fs"), data = data.Ehux, method = "REML")
summary(Ehux_modGS_sistar)
```

##train_ehux_modGS_sistar
```{r}
train_ehux_modGS_sistar <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(si_star, scientific_name, bs = "fs"), data = train.Ehux, method = "REML")
summary(train_ehux_modGS_sistar)
```

##test_ehux_modGS_sistar

```{r}
test_ehux_modGS_sistar <- train_ehux_modGS_sistar %>% predict(test.Ehux, type = "response")
summary(test_ehux_modGS_sistar)
```

RMSE of test data: 
```{r}
error_ehux_modGS_sistar <- test.Ehux$cells_per_litre - test_ehux_modGS_sistar
RMSE_ehux_modGS_sistar <- sqrt(mean(error_ehux_modGS_sistar^2))
print(RMSE_ehux_modGS_sistar)
```

prediction error rate:
```{r}
RMSE_ehux_modGS_sistar/mean(test.Ehux$cells_per_litre)
```

#Ehux_modGS_nistar

```{r}
Ehux_modGS_nistar <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp")  + te(longitude, latitude, bs = c("cc", "tp")) + s(ni_star, scientific_name, bs = "fs"), data = data.Ehux, method = "REML")
summary(Ehux_modGS_nistar)
```

##train_ehux_modGS_nistar
```{r}
train_ehux_modGS_nistar <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(ni_star, scientific_name, bs = "fs"), data = train.Ehux, method = "REML")
summary(train_ehux_modGS_nistar)
```

##test_ehux_modGS_nistar

```{r}
test_ehux_modGS_nistar <- train_ehux_modGS_nistar %>% predict(test.Ehux, type = "response")
summary(test_ehux_modGS_nistar)
```

RMSE of test data: 
```{r}
error_ehux_modGS_nistar <- test.Ehux$cells_per_litre - test_ehux_modGS_nistar
RMSE_ehux_modGS_nistar <- sqrt(mean(error_ehux_modGS_nistar^2))
print(RMSE_ehux_modGS_nistar)
```

prediction error rate:
```{r}
RMSE_ehux_modGS_nistar/mean(test.Ehux$cells_per_litre)
```

#Ehux_modGS_s

```{r}
Ehux_modGS_s <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(s_an, scientific_name, bs = "fs"), data = data.Ehux, method = "REML")
summary(Ehux_modGS_s)
```

##train_ehux_modGS_s
```{r}
train_ehux_modGS_s <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(s_an, scientific_name, bs = "fs"), data = train.Ehux, method = "REML")
summary(train_ehux_modGS_s)
```

##test_ehux_modGS_s

```{r}
test_ehux_modGS_s <- train_ehux_modGS_s %>% predict(test.Ehux, type = "response")
summary(test_ehux_modGS_s)
```

RMSE of test data: 
```{r}
error_ehux_modGS_s <- test.Ehux$cells_per_litre - test_ehux_modGS_s
RMSE_ehux_modGS_s <- sqrt(mean(error_ehux_modGS_s^2))
print(RMSE_ehux_modGS_s)
```

prediction error rate:
```{r}
RMSE_ehux_modGS_s/mean(test.Ehux$cells_per_litre)
```

#Ehux_modGS_t

```{r}
Ehux_modGS_t <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(t_an, scientific_name, bs = "fs"), data = data.Ehux, method = "REML")
summary(Ehux_modGS_t)
```

##train_ehux_modGS_t
```{r}
train_ehux_modGS_t <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(t_an, scientific_name, bs = "fs"), data = train.Ehux, method = "REML")
summary(train_ehux_modGS_t)
```

##test_ehux_modGS_t

```{r}
test_ehux_modGS_t <- train_ehux_modGS_t %>% predict(test.Ehux, type = "response")
summary(test_ehux_modGS_t)
```

RMSE of test data: 
```{r}
error_ehux_modGS_t <- test.Ehux$cells_per_litre - test_ehux_modGS_t
RMSE_ehux_modGS_t <- sqrt(mean(error_ehux_modGS_t^2))
print(RMSE_ehux_modGS_t)
```

prediction error rate:
```{r}
RMSE_ehux_modGS_t/mean(test.Ehux$cells_per_litre)
```

##MODEL GI - environmental variables + individual smoother scientific_name

#Ehux_modGI_p

```{r}
Ehux_modGS_p <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(p_an, by = scientific_name, bs = "fs", m = 1), data = data.Ehux, method = "REML")
summary(Ehux_modGS_p)
```

###train_ehux_modGI_p
```{r}
train_ehux_modGI_p <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(p_an, by = scientific_name, bs = "fs", m = 1), data = train.Ehux, method = "REML")
summary(train_ehux_modGI_p)
```

##test_ehux_modGI_p

```{r}
test_ehux_modGI_p <- train_ehux_modGI_p %>% predict(test.Ehux, type = "response")
summary(test_ehux_modGI_p)
```

RMSE of test data: 
```{r}
error_ehux_modGI_p <- test.Ehux$cells_per_litre - test_ehux_modGI_p
RMSE_ehux_modGI_p <- sqrt(mean(error_ehux_modGI_p^2))
print(RMSE_ehux_modGI_p)
```

prediction error rate:
```{r}
RMSE_ehux_modGI_p/mean(test.Ehux$cells_per_litre)
```

#Ehux_modGI_sistar

```{r}
Ehux_modGS_sistar <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(si_star, by = scientific_name, bs = "fs", m = 1), data = data.Ehux, method = "REML")
summary(Ehux_modGS_sistar)
```

###train_ehux_modGI_sistar
```{r}
train_ehux_modGI_sistar <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(si_star, by = scientific_name, bs = "fs", m = 1), data = train.Ehux, method = "REML")
summary(train_ehux_modGI_sistar)
```

##test_ehux_modGI_sistar

```{r}
test_ehux_modGI_sistar <- train_ehux_modGI_sistar %>% predict(test.Ehux, type = "response")
summary(test_ehux_modGI_sistar)
```

RMSE of test data: 
```{r}
error_ehux_modGI_sistar <- test.Ehux$cells_per_litre - test_ehux_modGI_sistar
RMSE_ehux_modGI_sistar <- sqrt(mean(error_ehux_modGI_sistar^2))
print(RMSE_ehux_modGI_sistar)
```

prediction error rate:
```{r}
RMSE_ehux_modGI_sistar/mean(test.Ehux$cells_per_litre)
```

#Ehux_modGI_nistar

```{r}
Ehux_modGI_nistar <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp")  + te(longitude, latitude, bs = c("cc", "tp")) + s(ni_star, by = scientific_name, bs = "fs", m = 1), data = data.Ehux, method = "REML")
summary(Ehux_modGI_nistar)
```

###train_ehux_modGI_nistar
```{r}
train_ehux_modGI_nistar <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(ni_star, by = scientific_name, bs = "fs", m = 1), data = train.Ehux, method = "REML")
summary(train_ehux_modGI_nistar)
```

##test_ehux_modGI_nistar

```{r}
test_ehux_modGI_nistar <- train_ehux_modGI_nistar %>% predict(test.Ehux, type = "response")
summary(test_ehux_modGI_nistar)
```

RMSE of test data: 
```{r}
error_ehux_modGI_nistar <- test.Ehux$cells_per_litre - test_ehux_modGI_nistar
RMSE_ehux_modGI_nistar <- sqrt(mean(error_ehux_modGI_nistar^2))
print(RMSE_ehux_modGI_nistar)
```

prediction error rate:
```{r}
RMSE_ehux_modGI_nistar/mean(test.Ehux$cells_per_litre)
```

#Ehux_modGI_s

```{r}
Ehux_modGI_s <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(s_an, by =  scientific_name, bs = "fs", m = 1), data = data.Ehux, method = "REML")
summary(Ehux_modGI_s)
```

###train_ehux_modGI_s
```{r}
train_ehux_modGI_s <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(s_an, by = scientific_name, bs = "fs", m = 1), data = train.Ehux, method = "REML")
summary(train_ehux_modGI_s)
```

##test_ehux_modGI_s

```{r}
test_ehux_modGI_s <- train_ehux_modGI_s %>% predict(test.Ehux, type = "response")
summary(test_ehux_modGI_s)
```

RMSE of test data: 
```{r}
error_ehux_modGI_s <- test.Ehux$cells_per_litre - test_ehux_modGI_s
RMSE_ehux_modGI_s <- sqrt(mean(error_ehux_modGI_s^2))
print(RMSE_ehux_modGI_s)
```

prediction error rate:
```{r}
RMSE_ehux_modGI_s/mean(test.Ehux$cells_per_litre)
```

#Ehux_modGI_t

```{r}
Ehux_modGI_t <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(t_an, by = scientific_name, bs = "fs", m = 1), data = data.Ehux, method = "REML")
summary(Ehux_modGI_t)
```


###train_ehux_modGI_t
```{r}
train_ehux_modGI_t <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(t_an, by = scientific_name, bs = "fs", m = 1), data = train.Ehux, method = "REML")
summary(train_ehux_modGI_t)
```

##test_ehux_modGI_t

```{r}
test_ehux_modGI_t <- train_ehux_modGI_t %>% predict(test.Ehux, type = "response")
summary(test_ehux_modGI_t)
```

RMSE of test data: 
```{r}
error_ehux_modGI_t <- test.Ehux$cells_per_litre - test_ehux_modGI_t
RMSE_ehux_modGI_t <- sqrt(mean(error_ehux_modGI_t^2))
print(RMSE_ehux_modGI_t)
```

prediction error rate:
```{r}
RMSE_ehux_modGI_t/mean(test.Ehux$cells_per_litre)
```


#AIC values
```{r}
AIC(train_ehux_modG, train_ehux_modGS_p, train_ehux_modGS_sistar, train_ehux_modGS_nistar, train_ehux_modGS_s, train_ehux_modGS_t, train_ehux_modGI_p, train_ehux_modGI_sistar, train_ehux_modGI_nistar, train_ehux_modGI_s, train_ehux_modGI_t)
```












