---
title: "Ehux models"
author: "Seoyeon Park B00768397"
date: "10/28/2021"
output:
  word_document: default
  html_document: default
---

#Scientific name is only one factor level so gam returns an error - need to remove s(scientific_name, bs = "re")
`
```{r}
load("data_subsets.RData")
```

```{r}
library(mgcv)
```


##MODEL G

#Ehux_modG

```{r}
Ehux_modG <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc","tp")) , data = data.Ehux)
summary(Ehux_modG)
```

##MODEL GS - environmental variables + group smoother scientific_name

#Ehux_modGS_p

```{r}
Ehux_modGS_p <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(p_an, scientific_name, bs = "fs"), data = data.Ehux, method = "REML")
summary(Ehux_modGS_p)
```


#Ehux_modGS_sistar

```{r}
Ehux_modGS_sistar <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(si_star, scientific_name, bs = "fs"), data = data.Ehux, method = "REML")
summary(Ehux_modGS_sistar)
```

#Ehux_modGS_nistar

```{r}
Ehux_modGS_nistar <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp")  + te(longitude, latitude, bs = c("cc", "tp")) + s(ni_star, scientific_name, bs = "fs"), data = data.Ehux, method = "REML")
summary(Ehux_modGS_nistar)
```

#Ehux_modGS_s

```{r}
Ehux_modGS_s <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(s_an, scientific_name, bs = "fs"), data = data.Ehux, method = "REML")
summary(Ehux_modGS_s)
```

#Ehux_modGS_t

```{r}
Ehux_modGS_t <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(t_an, scientific_name, bs = "fs"), data = data.Ehux, method = "REML")
summary(Ehux_modGS_t)
```

##MODEL GI - environmental variables + individual smoother scientific_name

#Ehux_modGI_p

```{r}
Ehux_modGS_p <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(p_an, by = scientific_name, bs = "fs", m = 1), data = data.Ehux, method = "REML")
summary(Ehux_modGS_p)
```


#Ehux_modGI_sistar

```{r}
Ehux_modGS_sistar <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(si_star, by = scientific_name, bs = "fs", m = 1), data = data.Ehux, method = "REML")
summary(Ehux_modGS_sistar)
```

#Ehux_modGI_nistar

```{r}
Ehux_modGI_nistar <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp")  + te(longitude, latitude, bs = c("cc", "tp")) + s(ni_star, by = scientific_name, bs = "fs", m = 1), data = data.Ehux, method = "REML")
summary(Ehux_modGI_nistar)
```

#Ehux_modGI_s

```{r}
Ehux_modGI_s <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(s_an, by =  scientific_name, bs = "fs", m = 1), data = data.Ehux, method = "REML")
summary(Ehux_modGI_s)
```

#Ehux_modGI_t

```{r}
Ehux_modGI_t <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(t_an, by = scientific_name, bs = "fs", m = 1), data = data.Ehux, method = "REML")
summary(Ehux_modGI_t)
```



