---
title: "LASSO models"
author: "Seoyeon Park B00768397"
date: "11/24/2021"
output: html_document
---

#LASSO model full
 
```{r}
load("data_subsets.RData")
load("data_crossval.RData")
library(mgcv)
library(glmnet)
library(stats)
library(caret)
set.seed(101)
```
 
 
Organize data into matrix:
```{r}
y <- data.full$cells_per_litre
x <- data.matrix(data.full[, c("p_an", "si_star", "ni_star", "s_an", "t_an", "scientific_name", "longitude", "latitude")])
```

```{r}
cv_model_full <- cv.glmnet(x, y, alpha = 1)
best_lambda <- cv_model_full$lambda.min
best_lambda
plot(cv_model_full)
```

```{r}
best_model_full <- glmnet(x, y, alpha = 1, lambda = best_lambda)
coef(best_model_full)
```

No predictors completely shrunk to 0. 

##train_lasso_full
```{r}
y <- train.full$cells_per_litre
x <- data.matrix(train.full[, c("p_an", "si_star", "ni_star", "s_an", "t_an", "scientific_name", "longitude", "latitude")])
```

```{r}
cv_model_train_full <- cv.glmnet(x, y, alpha = 1)
best_lambda <- cv_model_train_full$lambda.min
best_lambda
plot(cv_model_train_full)
```

```{r}
best_model_lasso_full <- glmnet(x, y, alpha = 1, lambda = best_lambda)
coef(best_model_lasso_full)
```

salinity shrunk to 0 

```{r}
sqrt(mean((y - predict(best_model_lasso_full, newx = x, type = "response"))^2))
```



##test_lasso_full

```{r}
x <- data.matrix(test.full[, c("p_an", "si_star", "ni_star", "s_an", "t_an", "scientific_name", "longitude", "latitude")])
```

```{r}
test_lasso_full <- predict(best_model_lasso_full, newx = x, type = "response")
```

```{r}
summary(test_lasso_full)
```

RMSE of test data: 
```{r}
error_lasso_full <- test.full$cells_per_litre - test_lasso_full
RMSE_lasso_full <- sqrt(mean(error_lasso_full^2))
print(RMSE_lasso_full)
```

prediction error rate:
```{r}
RMSE_lasso_full/mean(test.full$cells_per_litre)
```

#LASSO model coccolithus
 
Organize data into matrix:
```{r}
y <- data.Coccolithus$cells_per_litre
x <- data.matrix(data.Coccolithus[, c("p_an", "si_star", "ni_star", "s_an", "t_an", "scientific_name", "longitude", "latitude")])
```

```{r}
cv_model_coccolithus <- cv.glmnet(x, y, alpha = 1)
best_lambda <- cv_model_coccolithus$lambda.min
best_lambda
plot(cv_model_coccolithus)
```

```{r}
best_model_coccolithus <- glmnet(x, y, alpha = 1, lambda = best_lambda)
coef(best_model_coccolithus)
```

##train_lasso_coccolithus
```{r}
y <- train.Coccolithus$cells_per_litre
x <- data.matrix(train.Coccolithus[, c("p_an", "si_star", "ni_star", "s_an", "t_an", "scientific_name", "longitude", "latitude")])
```

```{r}
cv_model_train_coccolithus <- cv.glmnet(x, y, alpha = 1)
best_lambda <- cv_model_train_coccolithus$lambda.min
best_lambda
plot(cv_model_train_coccolithus)
```

```{r}
best_model_lasso_coccolithus <- glmnet(x, y, alpha = 1, lambda = best_lambda)
coef(best_model_lasso_coccolithus)
```

salinity shrunk to 0

```{r}
sqrt(mean((y - predict(best_model_lasso_coccolithus, newx = x, type = "response"))^2))
```

##test_lasso_coccolithus

```{r}
x <- data.matrix(test.Coccolithus[, c("p_an", "si_star", "ni_star", "s_an", "t_an", "scientific_name", "longitude", "latitude")])
```

```{r}
test_lasso_coccolithus <- predict(best_model_lasso_coccolithus, newx = x, type = "response")
```

```{r}
summary(test_lasso_coccolithus)
```

RMSE of test data: 
```{r}
error_lasso_coccolithus <- test.Coccolithus$cells_per_litre - test_lasso_coccolithus
RMSE_lasso_coccolithus <- sqrt(mean(error_lasso_coccolithus^2))
print(RMSE_lasso_coccolithus)
```

prediction error rate:
```{r}
RMSE_lasso_coccolithus/mean(test.full$cells_per_litre)
```

#LASSO model syracosphaera
 
Organize data into matrix:
```{r}
y <- data.Syracosphaera$cells_per_litre
x <- data.matrix(data.Syracosphaera[, c("p_an", "si_star", "ni_star", "s_an", "t_an", "scientific_name", "longitude", "latitude")])
```

```{r}
cv_model_syracosphaera <- cv.glmnet(x, y, alpha = 1)
best_lambda <- cv_model_syracosphaera$lambda.min
best_lambda
plot(cv_model_syracosphaera)
```

```{r}
best_model_syracosphaera <- glmnet(x, y, alpha = 1, lambda = best_lambda)
coef(best_model_syracosphaera)
```

None of the coefficients shrink to 0. 

##train_lasso_syracosphaera
```{r}
y <- train.Syracosphaera$cells_per_litre
x <- data.matrix(train.Syracosphaera[, c("p_an", "si_star", "ni_star", "s_an", "t_an", "scientific_name", "longitude", "latitude")])
```

```{r}
cv_model_train_syracosphaera <- cv.glmnet(x, y, alpha = 1)
best_lambda <- cv_model_train_syracosphaera$lambda.min
best_lambda
plot(cv_model_train_syracosphaera)
```

```{r}
best_model_lasso_syracosphaera <- glmnet(x, y, alpha = 1, lambda = best_lambda)
coef(best_model_lasso_syracosphaera)
```

```{r}
sqrt(mean((y - predict(best_model_lasso_syracosphaera, newx = x, type = "response"))^2))
```

##test_lasso_syracosphaera

```{r}
x <- data.matrix(test.Syracosphaera[, c("p_an", "si_star", "ni_star", "s_an", "t_an", "scientific_name", "longitude", "latitude")])
```

```{r}
test_lasso_syracosphaera <- predict(best_model_lasso_syracosphaera, newx = x, type = "response")
```

```{r}
summary(test_lasso_syracosphaera)
```

RMSE of test data: 
```{r}
error_lasso_syracosphaera <- test.Syracosphaera$cells_per_litre - test_lasso_syracosphaera
RMSE_lasso_syracosphaera <- sqrt(mean(error_lasso_syracosphaera^2))
print(RMSE_lasso_syracosphaera)
```

prediction error rate:
```{r}
RMSE_lasso_syracosphaera/mean(test.full$cells_per_litre)
```


=============================================================================================================================

#LASSO model Ehux
 
Organize data into matrix:
```{r}
y <- data.Ehux$cells_per_litre
x <- data.matrix(data.Ehux[, c("p_an", "si_star", "ni_star", "s_an", "t_an", "scientific_name", "longitude", "latitude")])
```

```{r}
cv_model_Ehux <- cv.glmnet(x, y, alpha = 1)
best_lambda <- cv_model_Ehux$lambda.min
best_lambda
plot(cv_model_Ehux)
```

```{r}
best_model_Ehux <- glmnet(x, y, alpha = 1, lambda = best_lambda)
coef(best_model_Ehux)
```

##train_lasso_Ehux
```{r}
y <- train.Ehux$cells_per_litre
x <- data.matrix(train.Ehux[, c("p_an", "si_star", "ni_star", "s_an", "t_an", "scientific_name", "longitude", "latitude")])
```

```{r}
cv_model_train_Ehux <- cv.glmnet(x, y, alpha = 1)
best_lambda <- cv_model_train_Ehux$lambda.min
best_lambda
plot(cv_model_train_Ehux)
```

```{r}
best_model_lasso_Ehux <- glmnet(x, y, alpha = 1, lambda = best_lambda)
coef(best_model_lasso_Ehux)
```

scientific name shrunk to 0 

```{r}
sqrt(mean((y - predict(best_model_lasso_Ehux, newx = x, type = "response"))^2))
```

##test_lasso_Ehux

```{r}
x <- data.matrix(test.Ehux[, c("p_an", "si_star", "ni_star", "s_an", "t_an", "scientific_name", "longitude", "latitude")])
```

```{r}
test_lasso_Ehux <- predict(best_model_lasso_Ehux, newx = x, type = "response")
```

```{r}
summary(test_lasso_Ehux)
```

RMSE of test data: 
```{r}
error_lasso_Ehux <- test.Ehux$cells_per_litre - test_lasso_Ehux
RMSE_lasso_Ehux <- sqrt(mean(error_lasso_Ehux^2))
print(RMSE_lasso_Ehux)
```

prediction error rate:
```{r}
RMSE_lasso_Ehux/mean(test.Ehux$cells_per_litre)
```

#LASSO model Cpel
 
Organize data into matrix:
```{r}
y <- data.Cpel$cells_per_litre
x <- data.matrix(data.Cpel[, c("p_an", "si_star", "ni_star", "s_an", "t_an", "scientific_name", "longitude", "latitude")])
```

```{r}
cv_model_Cpel <- cv.glmnet(x, y, alpha = 1)
best_lambda <- cv_model_Cpel$lambda.min
best_lambda
plot(cv_model_Cpel)
```

```{r}
best_model_Cpel <- glmnet(x, y, alpha = 1, lambda = best_lambda)
coef(best_model_Cpel)
```

##train_lasso_Cpel
```{r}
y <- train.Cpel$cells_per_litre
x <- data.matrix(train.Cpel[, c("p_an", "si_star", "ni_star", "s_an", "t_an", "scientific_name", "longitude", "latitude")])
```

```{r}
cv_model_train_Cpel <- cv.glmnet(x, y, alpha = 1)
best_lambda <- cv_model_train_Cpel$lambda.min
best_lambda
plot(cv_model_train_Cpel)
```

```{r}
best_model_lasso_Cpel <- glmnet(x, y, alpha = 1, lambda = best_lambda)
coef(best_model_lasso_Cpel)
```

```{r}
sqrt(mean((y - predict(best_model_lasso_Cpel, newx = x, type = "response"))^2))
```


##test_lasso_Cpel

```{r}
x <- data.matrix(test.Cpel[, c("p_an", "si_star", "ni_star", "s_an", "t_an", "scientific_name", "longitude", "latitude")])
```

```{r}
test_lasso_Cpel <- predict(best_model_lasso_Cpel, newx = x, type = "response")
```

```{r}
summary(test_lasso_Cpel)
```

RMSE of test data: 
```{r}
error_lasso_Cpel <- test.Cpel$cells_per_litre - test_lasso_Cpel
RMSE_lasso_Cpel<- sqrt(mean(error_lasso_Cpel^2))
print(RMSE_lasso_Cpel)
```

prediction error rate:
```{r}
RMSE_lasso_Cpel/mean(test.full$cells_per_litre)
```

#LASSO model Spul

Organize data into matrix:
```{r}
y <- data.Spul$cells_per_litre
x <- data.matrix(data.Spul[, c("p_an", "si_star", "ni_star", "s_an", "t_an", "scientific_name", "longitude", "latitude")])
```

```{r}
cv_model_Spul <- cv.glmnet(x, y, alpha = 1)
best_lambda <- cv_model_Spul$lambda.min
best_lambda
plot(cv_model_Spul)
```

```{r}
best_model_Spul <- glmnet(x, y, alpha = 1, lambda = best_lambda)
coef(best_model_Spul)
```

##train_lasso_Spul
```{r}
y <- train.Spul$cells_per_litre
x <- data.matrix(train.Spul[, c("p_an", "si_star", "ni_star", "s_an", "t_an", "scientific_name", "longitude", "latitude")])
```

```{r}
cv_model_train_Spul<- cv.glmnet(x, y, alpha = 1)
best_lambda <- cv_model_train_Spul$lambda.min
best_lambda
plot(cv_model_train_Spul)
```

```{r}
best_model_lasso_Spul <- glmnet(x, y, alpha = 1, lambda = best_lambda)
coef(best_model_lasso_Spul)
```

```{r}
sqrt(mean((y - predict(best_model_lasso_Spul, newx = x, type = "response"))^2))
```

##test_lasso_Spul

```{r}
x <- data.matrix(test.Spul[, c("p_an", "si_star", "ni_star", "s_an", "t_an", "scientific_name", "longitude", "latitude")])
```

```{r}
test_lasso_Spul <- predict(best_model_lasso_Spul, newx = x, type = "response")
```

```{r}
summary(test_lasso_Spul)
```

RMSE of test data: 
```{r}
error_lasso_Spul <- test.Spul$cells_per_litre - test_lasso_Spul
RMSE_lasso_Spul <- sqrt(mean(error_lasso_Spul^2))
print(RMSE_lasso_Spul)
```

prediction error rate:
```{r}
RMSE_lasso_Spul/mean(test.full$cells_per_litre)
```














