---
title: "Spul models"
author: "Seoyeon Park B00768397"
date: "10/28/2021"
output:
  word_document: default
  html_document: default
---

Syracosphaera pulchra

#Scientific name is only one factor level so gam returns an error - need to remove s(scientific_name, bs = "re")
`
```{r}
load("data_subsets.RData")
```

```{r}
library(mgcv)
library(glmnet)
library(stats)
library(caret)
set.seed(101)
```


##MODEL G

#Spul_modG

```{r}
Spul_modG <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc","tp")) , data = data.Spul)
summary(Spul_modG)
```

###train_spul_modG

```{r}
train_spul_modG <-  gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc","tp")) , data = train.Spul)
summary(train_spul_modG)
```

```{r}
sqrt(mean((train.Spul$cells_per_litre - predict(train_spul_modG, type = "response"))^2))
```

```{r}
plot(train_spul_modG, shade = TRUE, pages = 1, scale = 0, seWithMean = TRUE)
```

###test_spul_modG

```{r}
test_spul_modG <- train_spul_modG %>% predict(test.Spul, type = "response")
summary(test_spul_modG)
```



RMSE of test data: 
```{r}
error_spul_modG <- test.Spul$cells_per_litre - test_spul_modG
RMSE_spul_modG <- sqrt(mean(error_spul_modG^2))
print(RMSE_spul_modG)
```

prediction error rate:
```{r}
RMSE_spul_modG/mean(test.Spul$cells_per_litre)
```

Actual vs predicted: 

##MODEL GS - environmental variables + group smoother scientific_name

#Spul_modGS_p

```{r}
Spul_modGS_p <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(p_an, scientific_name, bs = "fs"), data = data.Spul, method = "REML")
summary(Spul_modGS_p)
```

##train_spul_modGS_p
```{r}
train_spul_modGS_p <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(p_an, scientific_name, bs = "fs"), data = train.Spul, method = "REML")
summary(train_spul_modGS_p)
```

```{r}
sqrt(mean((train.Spul$cells_per_litre - predict(train_spul_modGS_p, type = "response"))^2))
```

```{r}
plot(train_spul_modGS_p, shade = TRUE, pages = 1, scale = 0, seWithMean = TRUE)
```

###test_spul_modGS_p

```{r}
test_spul_modGS_p <- train_spul_modGS_p %>% predict(test.Spul, type = "response")
summary(test_spul_modGS_p)
```

RMSE of test data: 
```{r}
error_spul_modGS_p <- test.Spul$cells_per_litre - test_spul_modGS_p
RMSE_spul_modGS_p <- sqrt(mean(error_spul_modGS_p^2))
print(RMSE_spul_modGS_p)
```

prediction error rate:
```{r}
RMSE_spul_modGS_p/mean(test.Spul$cells_per_litre)
```


#Spul_modGS_sistar

```{r}
Spul_modGS_sistar <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(si_star, scientific_name, bs = "fs"), data = data.Spul, method = "REML")
summary(Spul_modGS_sistar)
```

##train_spul_modGS_sistar
```{r}
train_spul_modGS_sistar <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(si_star, scientific_name, bs = "fs"), data = train.Spul, method = "REML")
summary(train_spul_modGS_sistar)
```

```{r}
sqrt(mean((train.Spul$cells_per_litre - predict(train_spul_modGS_sistar, type = "response"))^2))
```

```{r}
plot(train_spul_modGS_sistar, shade = TRUE, pages = 1, scale = 0, seWithMean = TRUE)
```

###test_spul_modGS_sistar

```{r}
test_spul_modGS_sistar <- train_spul_modGS_sistar %>% predict(test.Spul, type = "response")
summary(test_spul_modGS_sistar)
```

RMSE of test data: 
```{r}
error_spul_modGS_sistar <- test.Spul$cells_per_litre - test_spul_modGS_sistar
RMSE_spul_modGS_sistar <- sqrt(mean(error_spul_modGS_sistar^2))
print(RMSE_spul_modGS_sistar)
```

prediction error rate:
```{r}
RMSE_spul_modGS_sistar/mean(test.Spul$cells_per_litre)
```

#Spul_modGS_nistar

```{r}
Spul_modGS_nistar <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(ni_star, scientific_name, bs = "fs"), data = data.Spul, method = "REML")
summary(Spul_modGS_nistar)
```

##train_spul_modGS_nistar
```{r}
train_spul_modGS_nistar <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(ni_star, scientific_name, bs = "fs"), data = train.Spul, method = "REML")
summary(train_spul_modGS_nistar)
```

```{r}
sqrt(mean((train.Spul$cells_per_litre - predict(train_spul_modGS_nistar, type = "response"))^2))
```

```{r}
plot(train_spul_modGS_nistar, shade = TRUE, pages = 1, scale = 0, seWithMean = TRUE)
```

###test_spul_modGS_nistar

```{r}
test_spul_modGS_nistar <- train_spul_modGS_nistar %>% predict(test.Spul, type = "response")
summary(test_spul_modGS_nistar)
```

RMSE of test data: 
```{r}
error_spul_modGS_nistar <- test.Spul$cells_per_litre - test_spul_modGS_nistar
RMSE_spul_modGS_nistar <- sqrt(mean(error_spul_modGS_nistar^2))
print(RMSE_spul_modGS_nistar)
```

prediction error rate:
```{r}
RMSE_spul_modGS_sistar/mean(test.Spul$cells_per_litre)
```

#Spul_modGS_s

```{r}
Spul_modGS_s <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(s_an, scientific_name, bs = "fs"), data = data.Spul, method = "REML")
summary(Spul_modGS_s)
```

##train_spul_modGS_s
```{r}
train_spul_modGS_s <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(s_an, scientific_name, bs = "fs"), data = train.Spul, method = "REML")
summary(train_spul_modGS_s)
```

```{r}
sqrt(mean((train.Spul$cells_per_litre - predict(train_spul_modGS_s, type = "response"))^2))
```

```{r}
plot(train_spul_modGS_s, shade = TRUE, pages = 1, scale = 0, seWithMean = TRUE)
```

###test_spul_modGS_s

```{r}
test_spul_modGS_s <- train_spul_modGS_s %>% predict(test.Spul, type = "response")
summary(test_spul_modGS_s)
```

RMSE of test data: 
```{r}
error_spul_modGS_s <- test.Spul$cells_per_litre - test_spul_modGS_p
RMSE_spul_modGS_s <- sqrt(mean(error_spul_modGS_p^2))
print(RMSE_spul_modGS_p)
```

prediction error rate:
```{r}
RMSE_spul_modGS_p/mean(test.Spul$cells_per_litre)
```

#Spul_modGS_t

```{r}
Spul_modGS_t <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(t_an, scientific_name, bs = "fs"), data = data.Spul, method = "REML")
summary(Spul_modGS_t)
```

##train_spul_modGS_t
```{r}
train_spul_modGS_t <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(t_an, scientific_name, bs = "fs"), data = train.Spul, method = "REML")
summary(train_spul_modGS_t)
```

```{r}
sqrt(mean((train.Spul$cells_per_litre - predict(train_spul_modGS_t, type = "response"))^2))
```

```{r}
plot(train_spul_modGS_t, shade = TRUE, pages = 1, scale = 0, seWithMean = TRUE)
```

###test_spul_modGS_t

```{r}
test_spul_modGS_t <- train_spul_modGS_t %>% predict(test.Spul, type = "response")
summary(test_spul_modGS_t)
```

RMSE of test data: 
```{r}
error_spul_modGS_t <- test.Spul$cells_per_litre - test_spul_modGS_t
RMSE_spul_modGS_t <- sqrt(mean(error_spul_modGS_t^2))
print(RMSE_spul_modGS_t)
```

prediction error rate:
```{r}
RMSE_spul_modGS_t/mean(test.Spul$cells_per_litre)
```

##MODEL GI - environmental variables + individual smoother scientific_name

#Spul_modGI_p

```{r}
Spul_modGI_p <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(p_an, by = scientific_name, bs = "fs", m = 1), data = data.Spul, method = "REML")
summary(Spul_modGI_p)
```

###train_spul_modGI_p
```{r}
train_spul_modGI_p <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(p_an, by = scientific_name, bs = "fs", m = 1), data = train.Spul, method = "REML")
summary(train_spul_modGI_p)
```

```{r}
sqrt(mean((train.Spul$cells_per_litre - predict(train_spul_modGI_p, type = "response"))^2))
```

```{r}
plot(train_spul_modGI_p, shade = TRUE, pages = 1, scale = 0, seWithMean = TRUE)
```

###test_spul_modGI_p

```{r}
test_spul_modGI_p <- train_spul_modGI_p %>% predict(test.Spul, type = "response")
summary(test_spul_modGI_p)
```

RMSE of test data: 
```{r}
error_spul_modGI_p <- test.Spul$cells_per_litre - test_spul_modGI_p
RMSE_spul_modGI_p <- sqrt(mean(error_spul_modGI_p^2))
print(RMSE_spul_modGI_p)
```

prediction error rate:
```{r}
RMSE_spul_modGI_p/mean(test.Spul$cells_per_litre)
```

#Spul_modGI_sistar

```{r}
Spul_modGI_sistar <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(si_star, by =  scientific_name, bs = "fs", m = 1), data = data.Spul, method = "REML")
summary(Spul_modGI_sistar)
```

###train_spul_modGI_sistar
```{r}
train_spul_modGI_sistar <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(si_star, by = scientific_name, bs = "fs", m = 1), data = train.Spul, method = "REML")
summary(train_spul_modGI_sistar)
```

```{r}
sqrt(mean((train.Spul$cells_per_litre - predict(train_spul_modGI_sistar, type = "response"))^2))
```

```{r}
plot(train_spul_modGI_sistar, shade = TRUE, pages = 1, scale = 0, seWithMean = TRUE)
```

###test_spul_modGI_sistar

```{r}
test_spul_modGI_sistar <- train_spul_modGI_sistar %>% predict(test.Spul, type = "response")
summary(test_spul_modGI_sistar)
```

RMSE of test data: 
```{r}
error_spul_modGI_sistar <- test.Spul$cells_per_litre - test_spul_modGI_sistar
RMSE_spul_modGI_sistar <- sqrt(mean(error_spul_modGI_sistar^2))
print(RMSE_spul_modGI_sistar)
```

prediction error rate:
```{r}
RMSE_spul_modGI_sistar/mean(test.Spul$cells_per_litre)
```

#Spul_modGI_nistar

```{r}
Spul_modGI_nistar <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(ni_star, by = scientific_name, bs = "fs", m = 1), data = data.Spul, method = "REML")
summary(Spul_modGI_nistar)
```

###train_spul_modGI_nistar
```{r}
train_spul_modGI_nistar <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(ni_star, by = scientific_name, bs = "fs", m = 1), data = train.Spul, method = "REML")
summary(train_spul_modGI_nistar)
```

```{r}
sqrt(mean((train.Spul$cells_per_litre - predict(train_spul_modGI_nistar, type = "response"))^2))
```

```{r}
plot(train_spul_modGI_nistar, shade = TRUE, pages = 1, scale = 0, seWithMean = TRUE)
```

###test_spul_modGI_nistar

```{r}
test_spul_modGI_nistar <- train_spul_modGI_nistar %>% predict(test.Spul, type = "response")
summary(test_spul_modGI_nistar)
```

RMSE of test data: 
```{r}
error_spul_modGI_nistar <- test.Spul$cells_per_litre - test_spul_modGI_nistar
RMSE_spul_modGI_nistar <- sqrt(mean(error_spul_modGI_nistar^2))
print(RMSE_spul_modGI_nistar)
```

prediction error rate:
```{r}
RMSE_spul_modGI_nistar/mean(test.Spul$cells_per_litre)
```

#Spul_modGI_s

```{r}
Spul_modGI_s <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(s_an, by = scientific_name, bs = "fs", m = 1), data = data.Spul, method = "REML")
summary(Spul_modGI_s)
```

###train_spul_modGI_s
```{r}
train_spul_modGI_s <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(s_an, by = scientific_name, bs = "fs", m = 1), data = train.Spul, method = "REML")
summary(train_spul_modGI_s)
```

```{r}
sqrt(mean((train.Spul$cells_per_litre - predict(train_spul_modGI_s, type = "response"))^2))
```

```{r}
plot(train_spul_modGI_s, shade = TRUE, pages = 1, scale = 0, seWithMean = TRUE)
```

###test_spul_modGI_s

```{r}
test_spul_modGI_s <- train_spul_modGI_s %>% predict(test.Spul, type = "response")
summary(test_spul_modGI_s)
```

RMSE of test data: 
```{r}
error_spul_modGI_s <- test.Spul$cells_per_litre - test_spul_modGI_s
RMSE_spul_modGI_s <- sqrt(mean(error_spul_modGI_s^2))
print(RMSE_spul_modGI_s)
```

prediction error rate:
```{r}
RMSE_spul_modGI_s/mean(test.Spul$cells_per_litre)
```

#Spul_modGI_t

```{r}
Spul_modGI_t <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(t_an, by = scientific_name, bs = "fs", m = 1), data = data.Spul, method = "REML")
summary(Spul_modGI_t)
```

###train_spul_modGI_t
```{r}
train_spul_modGI_t <- gam(cells_per_litre ~ s(p_an, bs = "tp") + s(si_star, bs = "tp") + s(ni_star, bs = "tp") + s(s_an, bs = "tp") + s(t_an, bs = "tp") + te(longitude, latitude, bs = c("cc", "tp")) + s(t_an, by = scientific_name, bs = "fs", m = 1), data = train.Spul, method = "REML")
summary(train_spul_modGI_t)
```

```{r}
sqrt(mean((train.Spul$cells_per_litre - predict(train_spul_modGI_t, type = "response"))^2))
```

```{r}
plot(train_spul_modGI_t, shade = TRUE, pages = 1, scale = 0, seWithMean = TRUE)
```

###test_spul_modGI_t

```{r}
test_spul_modGI_t <- train_spul_modGI_t %>% predict(test.Spul, type = "response")
summary(test_spul_modGI_t)
```

RMSE of test data: 
```{r}
error_spul_modGI_t <- test.Spul$cells_per_litre - test_spul_modGI_t
RMSE_spul_modGI_t <- sqrt(mean(error_spul_modGI_t^2))
print(RMSE_spul_modGI_t)
```

prediction error rate:
```{r}
RMSE_spul_modGI_t/mean(test.Spul$cells_per_litre)
```

#AIC values
```{r}
AIC(train_spul_modG, train_spul_modGS_p, train_spul_modGS_sistar, train_spul_modGS_nistar, train_spul_modGS_s, train_spul_modGS_t, train_spul_modGI_p, train_spul_modGI_sistar, train_spul_modGI_nistar, train_spul_modGI_s, train_spul_modGI_t)
```












